---
title: "Econ144 Project1"
author: "Aditya Gorla, Vincent Chenneveau, Georges Ip, Filip Karasovsky"
date: "4/24/2019"
output: html_document
---

##\n\n

```{r setup, include=FALSE}
rm(list=ls(all=TRUE))

library(car)
require(stats)
require(stats4)
library(MASS)
library(knitr)
library(tseries)
library(forecast)
library(moments)
library("readxl")
library(dynlm)
library(zoo)
library("TTR")
library(rlist)
library(tstools)
library(datasets)
library(fpp2)
library(seasonal)
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
Over the past few years developments in electric car technology and batteries have sky-rocketed.Tesla is a company that aims to be at the forefront of these technological advancements.Their first mover advantage into many industries such as SpaceX and the Boring Company has made them the headlines.Tesla has gained noteriety for its erratic CEO who has smoked weed on live broadcast and shot his car into space. Will Tesla maintain its status in the eyes of the public or will an overpaid CEO leave to much room for entrants and lose their competitive advantage? The public loves Tesla but Wall Street does not love an unpredictable CEO. Investigating the stock price of Tesla will give us an idea of where its headed in the future.




##\n\n

We chose to model and forecast the weekly adjusted closing price of Tesla stock over a 9 year period. We chose weekly data over daily data to avoid short term noise in higher frequency trading, as well as to make accurate trend and seasonality measures. 


##\n\n

## Results

#### Modeling and Forecasting Trend

```{r results='hold',message=FALSE}
tsla<-read.table("TSLA.csv", header=TRUE, sep=",") #Load the data into R
names(tsla)=c("date","open","high","low","close","adjClose","Volume") #name coloumns of table
attach(tsla)
tsla_ts<-ts(tsla$adjClose, start = 2010.5, frequency=52) #extract the adjusted close data
t<-seq(length=length(tsla_ts)) #create a sequence same length as data
t2<-t^2
t3<-t^3
t4<-t^4
t5<-t^5
ltsla_ts<- log(tsla_ts) #take the logo of price data

plot(tsla_ts, xlab="Year", ylab="Adjusted Closing Price (in $)", main="TSLA Stock")# Part A
tsdisplay(tsla_ts, main="TSLA Stock", xlab="Year", ylab="Adjusted Closing Price (in $)")

```

1.b) The plot shows that the stock prices are increasing over time. This indicates that the mean is time variant. In addition, The variance changes across time, with variance being relatively flat initially, before undergoing periods of fluctuating volatility. Hence, the data is not covariance stationary.

##\n\n

1.c)The ACF shows that the data follows a correlated process, although it decays over time. This indicates that realizations from periods far into the past have less correlation with present realizations.The PACF shows that the spikes are mostly not significant, with only a few spikes exceeding the confidence intervals. Unlike the ACF, the PACF does not seem to decay over time. The ACF and PACF suggest that the series is autoregressive; the ACF gradually decays to zero while the pacf exhibits several major lag spikes, namely at lag 1 and lag 10. We can conclude that an AR model would be best suited for forecasting TESLA stock prices.

##\n\n

#### Initial Model building
```{r results='hold',message=FALSE}
# Part d
lin_m<- dynlm(tsla_ts ~ t) # linear model
plot(tsla_ts,lwd=2, col=1, main="TSLA Stock Linear Model Fit", xlab="Year", ylab="Adjusted Closing Price (in $)")
lines(lin_m$fitted.values ,col=2,lwd=2)
legend('bottomright', legend=c("Actual Value", "Linear Model Fit"), col=1:2, lwd=2)

nl_m<- dynlm(ltsla_ts ~ t + t2 + t3)# non-linear model
plot(ltsla_ts,lwd=2, col="black", main="TSLA Stock Non-Linear Model Fit", xlab="Year", ylab="Log-Adjusted Closing Price (in $)")
lines(nl_m$fitted.values,col='blue3',lwd=2)
legend('bottomright', legend=c("Actual Value", "Non-Linear Model Fit"), col=c("black", "blue3"),lwd=2)
```

##\n\n

#### Initial Model Diagnostics and Selection
```{r results='hold',message=FALSE}
#Part e,f,g, and h
# linear model
plot(y=lin_m$residuals, x=lin_m$fitted.values,lwd=2, col=1, main="TSLA Linear Model", xlab="Fitted Values", ylab="Residuals")
truehist(lin_m$residuals,col='gray80',main="TSLA Linear Model", ylab="Fraction", xlab="Residuals")
summary(lin_m)


# non-linear model
plot(y=nl_m$residuals, x=nl_m$fitted.values,lwd=2, col=1, main="TSLA Non-Linear Model", xlab="Fitted Values", ylab="Residuals")
truehist(nl_m$residuals,col='gray80',main="TSLA Non-Linear Model", ylab="Fraction", xlab="Residuals")
summary(nl_m)

AIC(lin_m,nl_m)

BIC(lin_m,nl_m)
```

For the linear model, the plot of the residuals and the fitted values shows that there is clearly a lot of structure present in the residuals. The residuals seem to go through some seasonality. For the linear model, the plot of the histogram shows that the residuals are centered around 0 and are somewhat symmetric but is skewed. Hence, the normality assumption for the residuals is unlikely to hold.

##\n\n

For the non-linear model, the plot of the residuals and the fitted values shows that there is clearly some structure present in the residuals. The residuals seem to go through some seasonality as well.Compared to the linear model, the plot of the histogram shows that the residuals also centered around 0 but are less symmetric but is skewed. Hence, the normality assumption for the residuals is not to hold as well. However, the scale of the residuals for the non-linear model is significantly smaller compared to the


##\n\n

For the linear model, we have an adjusted R-squared of 0.8562, which means that a relatively high proportion of the stock prices are explained by t. 
The t-value for the coefficient of t also indicates that the estimated coefficient is statistically significant at a confidence level of 95%.
From the F-statistic, we reject the null hypothesis that the coefficient of all the variables is equal to 0.

##\n\n

For the non-linear model, we have an adjusted R-squared of 0.9006, which means that a higher proportion of the stock prices are explained by t, t2 and t3. 
The t-value for the coefficients of t, t2 and t3 also indicates that the estimated coefficient is statistically significant at a confidence level of 95%.
From the F-statistic, we reject the null hypothesis that the coefficient of all the variables is equal to 0.

##\n\n

For AIC, the linear model has a value of 4794, but the non linear model has a much lower value of 275. Hence, based on AIC, we pick the non-linear model. For BIC, the linear model has a value of 4806, but the non linear model has a much lower value of 296 Hence, based on BIC, we pick the non-linear model. Overall, the AIC/BIC suggests that the nonlinear model is better than the linear model with regards to forecasting strength. Hence, both models agree.

##\n\n

#### Non-Linear Model Forecast
We are going to do a 24-steps forecast using our Non-Linear model above. This should predict the weekly Tesla adjusted closing price for the next 6 months.
```{r results='hold',message=FALSE}
#Part i
plot(predict(nl_m$fitted.values, h=24) ,shadecols="oldstyle", main="TSLA Stock Non-Linear Model Forecast", xlab="Year", ylab="Log-Adjusted Closing Price (in $)", lwd=2, col=1, ylim=c(2.75, 6))
lines(ltsla_ts,lwd=2, col=2)
legend('bottomright', legend=c( "Model+Forecast","Actual Value"), col=1:2, lwd=2)
```

##\n\n

### Modeling and Forecasting Seasonality

#### Seasonal Model Building and Analysis

```{r results='hold',message=FALSE}
seas_m=tslm(tsla_ts~season+0)
summary(seas_m)
plot(seas_m$coefficients, xlab="Week", ylab="Co-effecient Value",main="Estimated Seasonal Factors", type="o",lwd=2)

#Some seasonal decompositions to better understand seasonality
ggsubseriesplot(tsla_ts ,ylab="Adjusted Closing Price (in $)", xlab="Week") +
  ggtitle("TSLA Stock(w/ ggsubseriesplot()")
ggseasonplot(tsla_ts ,ylab="Adjusted Closing Price (in $)", xlab="Week") +
  ggtitle("TSLA Stock(w/ ggsubseriesplot()")
stl(tsla_ts,s.window=52) %>% autoplot() +
  ggtitle("STL decomposition of Tesla Stock Prices")
```
"""

TODO: - Comfirm we did 2.a and 2.b correctly
"""
The interpretation of this coefficient plot heavily suggests a pattern in stock price growth; a face-value analysis shows there is growth during the first half of each year (weeks 0-26) which then falls in the second half of the year. A continuation of this plot, however, into the next year, would suggest that growth occurs between ~week 30 of year i and week 30 of year i+1. Overall, the plot suggests a clear seasonal component where prices rise over ~60 weeks before falling again.

##\n\n

#### The Full Model

```{r results='hold',message=FALSE}
#cut data
fmltsla_ts=window(ltsla_ts,2010,c(2019,7))
frltsla_ts=window(ltsla_ts,c(2019,8),c(2019,19))
fmt=t[1:449]
fmt2=t2[1:449]
fmt3=t3[1:449]
full_m<-tslm(fmltsla_ts~ fmt + fmt2 + fmt3 + season + 0)

#model performance statistics
plot(ltsla_ts,lwd=2, col=1, main="TSLA Stock Full Model Fit", xlab="Year", ylab="Log-Adjusted Closing Price (in $)")
lines(full_m$fitted.values ,col=2,lwd=2)
legend('bottomright', legend=c("Actual Value", "Full Model Fit"), col=1:2, lwd=2)
plot(y=full_m$residuals,x=full_m$fitted.values,ylab="Residuals",xlab="fitted values", main="TSLA Full Model")
truehist(full_m$residuals,col='gray80',main="TSLA Full Model", ylab="Fraction", xlab="Residuals")
summary(full_m)
AIC(nl_m, full_m)
BIC(nl_m, full_m)
```

"""

TODO: - do word answers for 2.b, 2.c and 2.d
      - if confidence intervals are missing below, Add them!

"""
The TSLA full model residuals still contain structure, demonstrating a periodic trend, roughly equivalent to a sine wave. This tells us that the nonlinear model combined with the seasonal component is not suffcient to fully explain the signal exhibited in the data, as some of it is still retained in the residuals. We can therefore assume that the full model is missing some cyclical component, such as an AR or MA process. The T tests and F test confirm the significance of all variables in the full model well beyond a standard alpha value of 5%, suggesting that the model integrity is overall sound. The histogram, however, of full model residuals are not unimodal and show some gravitation away from 0, pointing to some pattern that was omitted from the model.

##\n\n

<<<<<<< HEAD
The overall adjusted R-squared is 0.9952. This indicates that a high percentage of variation in Tesla stock prices are explained by our model. All the coefficients are statistically significant at the 95% confidence level. The F-statistic 1723 on 55 and 406 Degrees of freedom with a p-value much smaller than 0.05, allows us to reject the null that the overall model is not significant. The AIC and BIC of the full model pose an issue because both results are greater than AIC and BIC for the nonlinear trend alone which suggest our full model is weaker. Despite some evidence that the AIC and BIC point to the full model being weaker, these figures can be ignored due to the fact that the comparrison is between a non-linear model with only 3 variables against a full model with 52 extra seasonal dummies, coming out to a grand total of 55 estimators in the full model. The AIC and BIC can be dismissed in favor of both the plot of the residuals and the R squared value, which both point to the full model as being more robust than the linear model. 
=======
The overall adjusted R-squared is 0.9952. This indicates that a high percentage of variation in Tesla stock prices are explained by our model. All the coefficients are statistically significant at the 95% confidence level. The F-statistic 1723 on 55 and 406 Degrees of freedom with a p-value much smaller than 1, allows us to reject the null that the overall model is not significant.

##\n\n

#### The Full Model Forecast
```{r results='hold',message=FALSE}
plot(predict(full_m$fitted.values, h=24) ,shadecols="oldstyle", main="TSLA Stock Full Forecast", xlab="Year", ylab="Log-Adjusted Closing Price (in $)", lwd=2, col=1, ylim=c(2.75, 6))
lines(ltsla_ts,lwd=2, col=2)
legend('bottomright', legend=c( "Model+Forecast","Actual Value"), col=1:2, lwd=2)
legend(3.5,2019,legend=c(""))

```

##\n\n

### Conclusions and Future Work
"""
We used our full model to forecast 24 steps ahead. With our data set, this would translate to a prediction of 6 weeks ahead. Our model predicts that over the next 6 weeks, the Log prices of Tesla stock would decrease. However, if we look at how our model fares compared to the actual value, we see that our model departs from the model at certain periods of time. 

In order to improve the model in the future, we could try to fit an ARIMA model in order to create a better fit with our actual data. This could allow us to account for cycles in addition to trend and seasonality which we already have in our full model. Going back to the residuals plot of our full model, we see some significant structure in the residuals after accounting for trend and seasonality. This could be addressed if we incorporate cycles into our model in the future.
"""

##\n\n

### References
"""

TODO: - Fill in References here
      - make better comment for large chunks of code

"""