---
title: "Econ144 Project1"
author: "Aditya Gorla"
date: "4/24/2019"
output: html_document
---

##\n\n

```{r setup, include=FALSE}
rm(list=ls(all=TRUE))

library(car)
require(stats)
require(stats4)
library(MASS)
library(knitr)
library(tseries)
library(forecast)
library(moments)
library("readxl")
library(dynlm)
library(zoo)
library("TTR")
library(rlist)
library(tstools)
library(datasets)
library(fpp2)
library(seasonal)
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
"""

TODO: PUT INTRO HERE

"""

##\n\n

### Results

#### Initial Data Exploration

```{r results='hold',message=FALSE}
#Loading the data
tsla<-read.table("TSLA.csv", header=TRUE, sep=",")
names(tsla)=c("date","open","high","low","close","adjClose","Volume")
attach(tsla)
#we'll be working with adjusted closing price and the data is weekly
tsla_ts<-ts(tsla$adjClose, start = 2010.5, frequency=52)
t<-seq(length=length(tsla_ts))
t2<-t^2
t3<-t^3
t4<-t^4
t5<-t^5
ltsla_ts<- log(tsla_ts)

plot(tsla_ts, xlab="Year", ylab="Adjusted Closing Price (in $)", main="TSLA Stock")
tsdisplay(tsla_ts, main="TSLA Stock", xlab="Year", ylab="Adjusted Closing Price (in $)")

```
"""

TODO: write initial analysis based on above plot
      The ACF and PACF suggest that the series is autoregressive; the ACF gradually decays to zero while the pacf exhibits several major lag spikes, namely at lag 1 and lag 10. We can conclude that an AR model would be best suited for forecasting TESLA stock prices.

"""

##\n\n

#### Initial Model building
```{r results='hold',message=FALSE}
# linear model
lin_m<- dynlm(tsla_ts ~ t)
plot(tsla_ts,lwd=2, col=1, main="TSLA Stock Linear Model Fit", xlab="Year", ylab="Adjusted Closing Price (in $)")
lines(lin_m$fitted.values ,col=2,lwd=2)
legend('bottomright', legend=c("Actual Value", "Linear Model Fit"), col=1:2, lwd=2)

# non-linear model
nl_m<- dynlm(ltsla_ts ~ t + t2+ t3)
plot(ltsla_ts,lwd=2, col="black", main="TSLA Stock Non-Linear Model Fit", xlab="Year", ylab="log-Adjusted Closing Price (in $)")
lines(nl_m$fitted.values,col='blue3',lwd=2)
legend('bottomright', legend=c("Actual Value", "Non-Linear Model Fit"), col=c("black", "blue3"),lwd=2)
```

##\n\n

#### Initial Model Diagnostics and Selection
```{r results='hold',message=FALSE}

# linear model
plot(y=lin_m$residuals, x=lin_m$fitted.values,lwd=2, col=1, main="TSLA Linear Model", xlab="Fitted Values", ylab="Residuals")
truehist(lin_m$residuals,col='gray80',main="TSLA Linear Model", ylab="Fraction", xlab="Residuals")
summary(lin_m)


# non-linear model
plot(y=nl_m$residuals, x=nl_m$fitted.values,lwd=2, col=1, main="TSLA Non-Linear Model", xlab="Fitted Values", ylab="Residuals")
truehist(nl_m$residuals,col='gray80',main="TSLA Non-Linear Model", ylab="Fraction", xlab="Residuals")
summary(nl_m)

AIC(lin_m,nl_m)
BIC(lin_m,nl_m)
```
"""

TODO: Analyze the above plots and answer 1.e, 1.f, 1.g and 1.h

NOTE: the scale for the residulas are vastly differnt!

"""

```{r results='hold',message=FALSE}

```



```{r results='hold',message=FALSE}

```


